---
output: github_document
---

# APD: Average Proportional Distance for Item Analysis from Scales

The **APD** package provides functions to compute the **Average Proportional Distance**, a measure of internal consistency based on pairwise proportional differences between item scores.\
This approach focuses on the *average discrepancy* between item responses (in agreement with Sturman et al., 2009) and complements traditional reliability indices.

## Features

-   Compute **Average Proportional Distance (APD)** for item sets.
-   Supports **bootstrap confidence intervals** for APD.
-   Provides group comparisons based on confidence intervals of the difference (MOVER method).
-   Computes average inter–item correlation and group comparison.
-   Returns structured outputs for reproducible analyses.

## Installation

You can install the development version from GitHub:

```{r install, eval = FALSE}
# install.packages("devtools")
devtools::install_github("cmerinos/APD")
```

## Example

```{r example, eval = FALSE}
library(APD)

# toy data: 10 persons x 5 items
set.seed(123)
dat.example1 <- matrix(sample(1:5, 50, replace = TRUE), ncol = 5)

# compute APD
APD(dat.example1, ncat = 5, ci = TRUE, level = 0.95, B = 500)
```

## Package Structure

-   `APD()` – Main function to compute Average Proportional Distance.
-   `APDmg()` – Compute Average Proportional Distance for multiple groups.
-   `iia()` – Inter–item average.

## References

Sturman, D., Cribbie, R. A., & Flett, G. L. (2009).\
The average distance between item values: A novel approach for estimating internal consistency.\
*Educational and Psychological Measurement*, 69(6), 913–932. <https://doi.org/10.1177/0734282908330937>

Briggs, S.R. and Cheek, J.M. (1986).\
The role of factor analysis in the development and evaluation of personality scales.\
*Journal of Personality*, 54, 106–148. <https://doi.org/10.1111/j.1467-6494.1986.tb00391.x>

Clark, L. A., & Watson, D. (1995). \
Constructing validity: Basic issues in objective scale development.\
*Psychological Assessment*, 7}(3), 309–319. <https://doi.org/10.1037/1040-3590.7.3.309>

Piedmont, R.L. (2014). Inter-item correlations.\
In A.C. Michalos (Ed.), *Encyclopedia of Quality of Life and Well-Being Research*.\
Springer, Dordrecht. <https://doi.org/10.1007/978-94-007-0753-5_1493>

## License

This package is released under the **MIT License**.
